---
title: "Econ 294A Final"
author: "Richard Ronson"
date: "March 14, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(cache=TRUE)
```

```{r package-load, include=F,message=FALSE}
library(dplyr)
library(knitr)
library(stargazer)
library(RSQLite)
library(ggplot2)
library(nycflights13)
```

##Part A

```{r,echo=F,message=F,include=F}
db <- nycflights13_sqlite()
a<-left_join(
  tbl(db, "flights"),
  tbl(db, "weather"),
  by = c("year","month","day","hour"))%>% 
  collect() %>%
  mutate(canceled = is.na(arr_time))

a$month<-as.factor(a$month)
```

```{r,echo=F,warning = F,include=T}
A1.reg<-lm(dep_delay~temp+humid+wind_speed+precip+pressure+visib+wind_dir,a)
A2.reg<-lm(canceled~temp+humid+wind_speed+precip+pressure+visib+wind_dir,a)
```

```{r, warning=F,echo=F,fig.height = 3., fig.width=6, fig.cap = "Preciptations effect on Depature Delays"}

a.plot<-ggplot(data = sample_n(a, 10000),aes(x=precip,y=dep_delay))
a.plot+geom_point()+geom_line()
```


The plot above illustrates the most significant relationship of part A. This is between precipitation and depature delay.



```{r,echo=F,header=F,results='asis'}
stargazer(A1.reg,A2.reg,header=F,type='latex')
```


The regression results of Part A aimed to figure out the relationship between various 

weather conditions with both delayed and cancelled flights. 

All of the weather condtions were statistically significant at the 99% level in

determining delayed depatures. Likewise only a few weather conditons played no

signifcant role in leading to a cancelled flight. These conditions include: humidity, 

wind speed and precipitation. I would like to take note of the most significant statistic 
on table one which was the effect of precipitation on depature delay. This 

was that on average holding all else constant as precipitation increases by 1 unit 

depature delay will increase by 49.29 mins. This far and away was the greatest find of 

the regression results of part A both in statistical signifiance and magnitude.

(reference to table 1)

##Part B
```{r,echo=F,warning = F,include=F}
b1.reg<-lm(dep_delay~month+day+hour,a)
b2.reg<-lm(canceled~month+day+hour,a)
```

```{r, warning=F,echo=F,fig.height = 3., fig.width=6, fig.cap = "Time of Years Effect on Depature Delays"}

a.plot<-ggplot(data = sample_n(a, 10000),aes(x=month,y=dep_delay))
a.plot+geom_point()+geom_line()

```



The plot above illustrates the most significant relationship of part B. This is between time of year denoted by month and depature delay.



```{r,echo=F, results='asis'}
stargazer(b1.reg,b2.reg,header=F,type='latex')
```


The regression results of Part B strived to determine the relationship of time of

day, week and year between both delayed and cancelled flights. With each month relative 

to January as it is represented in the constant. Both month and hour are statistically

significant in determining delayed depatures.The day of the week was not statistically

significant in determining delayed depatures.In regards to months and cancellations all

but 4 months were statistically significant.These months include March, August, October,

and November. Also time of day and the day of the week were both statistically

significant.In terms of magniture and statistical signifcance no month illustrated such a

strong influence on depature delay as did july. On average holding all else constant if

you were to fly in july you can expect departure delays to increase by 11.557 minutes.

(reference to table 2)


##Part C
```{r,echo=F, message=F}
a$dest<-as.factor(a$dest)
```

```{r,echo=T,warning = F,include=T}
c1.reg<-lm(dep_delay~dest,a)
c2.reg<-lm(canceled~dest,a)

```

```{r,include=F,warning=F,echo=F}
summary(c1.reg)
summary(c2.reg)
#Much too large of a table to create so will just speak about the results of most revelant destinations.
```



In Part C I attempted to find relationships between various destinations with both

depature delays and cancelled flights. The number of destinations is very large so I

decided to dicuss the destinations with statistically signifcant relationships with

departure delays and cancellations (I have omitted this table from the pdf because it 

would have have taken up several pages in itself). I have instead depicted the models I

created in order to determine the relationships.The most statistically significant

relationships between destination and depature delays came from these destinations:

BHM,CAE,DSM,OKC,RIC,TUL,and TYS. These were all statistically significant at a level

greater than 99%. The most statistically significant relationships between destination

and cancelled flights came from the following destinations: ALB, BDL, BGR, BHM, BNA, BWI,

CAE, CHO, CHS, CLE, CMH, CVG, DAY, DCA, DSM, DTW, GRR,GSO, GSP, IAD, IND, JAC, LGA, 

MCI, MEM, MHT, MSP, OKC, ORD, ORF, PHL, PIT, PVD, RDU,RIC, SAV, SDF, SRQ, TUL and TYS. 

These destinations were all statistically significant at a level greater than 99%. In 

other words there were quite a bit more destinations that had a statistically signifcant 

relationship with cancelled flights.

(table ommitted)


##Part D
```{r,echo=F,warning = F,include=F}
db <- nycflights13_sqlite()


d<-left_join(
  tbl(db, "flights"),
  tbl(db, "planes"),
  by = "tailnum")%>% 
  collect() %>%
  mutate(canceled = is.na(arr_time))
```

```{r,echo=F,warning = F,include=T}
d1.reg<-lm(dep_delay~engines+seats+engine,d)
d2.reg<-lm(canceled~engines+seats+engine,d)
```

```{r, warning=F,echo=F,fig.height = 3.5, fig.width=6, fig.cap = "Time of Years Effect on Cancelled Flights"}

a.plot<-ggplot(data = sample_n(d, 10000),aes(x=seats,y=canceled))
a.plot+geom_point()+geom_line()

```



The plot above illustrates the most significant relationship of part D. This is between the number of seats on the plane and cancelled flights.



```{r,echo=F, results='asis'}
stargazer(d1.reg,d2.reg,header=F,type='latex')
```



Finally for Part D the regression model was designed to determine the effect of various

plane characteristics on both depature delays and cancelled flights. Both the number of

seats and engines were statistically significant in determining cancelled flights whereas 
only the number of seats were statistically significant in determining depature delays.

Overall the magnitude of the effects of the various plane characteristics are incredibly

small and I believe are negligible in relative cancellations or depature delays. Also 

Engine type was not statistically signifcant in determining either depature delays or

cancelled flights.

(reference to table 3)
